Promotors' reports on the first draft of the thesis document.

### 2020-04-07 ChapterRampInput/ChapterRampInput.tex
Final version before internal defense

\glsresetall

@@

The goal of any measurement is accurately estimating the input using the response of the sensor in the transient state.
Here we consider the case where the input varies during the measurement.
The simplest case is when the input changes with a constant rate, such as being described with an affine function.
The inputs that vary at a constant rate activate the sensor gradually. 
An affine input estimation method is needed to estimate the slope and the intercept parameters of the affine function.

@@

The effectiveness and uncertainty of the proposed methods are evaluated with simulation studies.
The uncertainty is assessed, for the adaptive data-driven method, using a Taylor series expansion of the EWRLS estimate, and
for the ML method, using the derivatives of the residual error obtained from the to-be-minimized cost function.

@@

This section describes the subspace method for solving the affine input estimation problem.
The procedure is motivated by the step input estimation method that is formulated as a structured errors-in-variables (EIV) problem and is solved using recursive least-squares (RLS).
The exponentially weighted recursive least-squares (EWRLS) is a generalization of RLS that allows for the extension of the estimation method to reconstruct the affine input.

@@

The EWRLS solution to the problem (\ref{eqn:min_ewrls}) is implemented to facilitate the online implementation of the affine input estimation method.
When $\lambda=1$, the EWRLS solution is identical to that of the RLS.
When $\lambda<1$, the older residuals are weighted with lower values than the residuals of recent observations.
In this way, the solution of the minimization problem depends more on newer data, and less on older data.

@@

The EWRLS solution to the problem (\ref{eqn:min_ewrls}) provides an affine input  estimation $\widehat{\mathbf{u}}$. 
The affine input parameters $a$ and $b$ are estimated, in a second step, by fitting $\widehat{\mathbf{u}}$ to a straight line using linear regression, as follows
\begin{equation} 
  \begin{bmatrix} 1 & 1  \\ \vdots & \vdots \\ T & 1 \end{bmatrix} 
  \begin{bmatrix} \widehat{a} \\ \widehat{b}  \end{bmatrix} = 
  \begin{bmatrix} \widehat{u}(1)-\tau \\ \vdots \\ \widehat{u}(T)-\tau \end{bmatrix}, \label{eqn:LS}
\end{equation}
where it is considered the existance of the offset $\tau$.

>>>>

%!TEX root = ..\Gus-thesis.tex
\glsresetall

@@

The goal of any measurement is accurately estimating the input using the response of the sensor in the transient regime.
Here we consider the case where the input varies during the measurement.
The simplest case is when the input changes with a constant rate, such as being described with an affine function.
The inputs that vary at a constant rate activate the sensor gradually. 
An affine input estimation method is needed to estimate the slope and the intercept parameters of the affine function.

@@

The effectiveness and uncertainty of the proposed methods are evaluated with simulation studies.
The uncertainty is assessed, for the adaptive data-driven method, using a Taylor series expansion of the exponentially weighted recursive least-squares estimate, and
for the ML method, using the derivatives of the residual error obtained from the to-be-minimized cost function.

@@

This section describes the subspace method for solving the affine input estimation problem.
The procedure is motivated by the step input estimation method that is formulated as a structured errors-in-variables (EIV) problem and is solved using recursive least-squares (RLS).
The exponentially weighted recursive least-squares is a generalization of RLS that allows for the extension of the estimation method to reconstruct the affine input.

@@

The exponentially weighted RLS solution to the problem (\ref{eqn:min_ewrls}) is implemented to facilitate the online implementation of the affine input estimation method.
When $\lambda=1$, the exponentially weighted RLS solution is identical to that of the RLS.
When $\lambda<1$, the older residuals are weighted with lower values than the residuals of recent observations.
In this way, the solution of the minimization problem depends more on newer data, and less on older data.

@@

The exponentially weighted RLS solution to the problem (\ref{eqn:min_ewrls}) provides an affine input  estimation $\widehat{\mathbf{u}}$. 
The affine input parameters $a$ and $b$ are estimated, in a second step, by fitting $\widehat{\mathbf{u}}$ to a straight line using linear regression, as follows
\begin{equation} 
  \begin{bmatrix} 1 & 1  \\ \vdots & \vdots \\ T & 1 \end{bmatrix} 
  \begin{bmatrix} \widehat{a} \\ \widehat{b}  \end{bmatrix} = 
  \begin{bmatrix} \widehat{u}(1)-\tau \\ \vdots \\ \widehat{u}(T)-\tau \end{bmatrix}, \label{eqn:LS}
\end{equation}
where it is considered the existance of the offset $\tau$.

### 2020-04-07 ChapterStatisticalAnalysis/ChapterStatisticalAnalysis.tex
Final version before internal defense

%!TEX root = ..\Thesis.tex

\glsresetall

@@

\begin{quote}
\vspace{-0.5cm}
\emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I. Pintelon, R., Csurcsia, P.Z., and Verbeke, D., "Bias and covariance of the least-squares estimate in a structured errors-in-variables problem", Computational Statistics & Data Analysis journal, Vol. 144, 2020, ISSN 0167-9473, doi:10.1016/j.csda.2019.106893. \nocite{QuintanaCSDA} }\vfill{}
\end{quote}

@@

The mathematical expectation operator is applied to the Taylor series expansion of the LS estimate (\ref{eqn:xhatexp}).
After removing the terms with negligible expected value, and considering the structure of matrix $\mathbf{K}$, the estimation bias (\ref{eqn:biasdef}) is approximated by 
\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{Q}^{-1} \left( \left( \mathbf{K}^\top \mathbf{B}_1 - \mathbf{B}_2 \right) x - \left( \mathbf{K}^\top \mathbf{B}_3 - \mathbf{B}_4 \right) \right) , \end{aligned} \label{eqn:biasE} \end{equation}
whereas, the estimation covariance (\ref{eqn:covdef}) is approximated by 
\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right) , \end{aligned} \label{eqn:varE} \end{equation}
where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \Big\}$, $\mathbf{P} = \mathbf{K} \mathbf{K}^\dagger$, and $\mathbf{P}_\perp = \mathbf{I} - \mathbf{P}$. 
The derivation of equations (\ref{eqn:biasE}) and (\ref{eqn:varE}) is described in Appendix 1.
The expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$ and $\mathbf{C}_2$ are obtained using the results of Lemma \ref{lem:lemma1}.

@@

\begin{comment}

\section{STATISTICAL ANALYSIS - Ramp input}

>>>>

%!TEX root = ..\Gus-thesis.tex
\glsresetall

@@

\begin{quote}
%\vspace{-0.5cm}
\emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I. Pintelon, R., Csurcsia, P.Z., and Verbeke, D., "Bias and covariance of the least-squares estimate in a structured errors-in-variables problem", Computational Statistics & Data Analysis journal, Vol. 144, 2020, ISSN 0167-9473, doi:10.1016/j.csda.2019.106893. \nocite{QuintanaCSDA} }\vfill{}
\end{quote}

@@

The mathematical expectation operator is applied to the Taylor series expansion of the LS estimate (\ref{eqn:xhatexp}).
After removing the terms with negligible expected value, and considering the structure of matrix $\mathbf{K}$, the estimation bias (\ref{eqn:biasdef}) is approximated by 
\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{Q}^{-1} \left( \left( \mathbf{K}^\top \mathbf{B}_1 - \mathbf{B}_2 \right) \mathbf{x} - \left( \mathbf{K}^\top \mathbf{B}_3 - \mathbf{B}_4 \right) \right) , \end{aligned} \label{eqn:biasE} \end{equation}
whereas, the estimation covariance (\ref{eqn:covdef}) is approximated by 
\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \left( \sigma_{\bm{\epsilon}}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right) , \end{aligned} \label{eqn:varE} \end{equation}
where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \Big\}$, $\mathbf{P} = \mathbf{K} \mathbf{K}^\dagger$, and $\mathbf{P}_\perp = \mathbf{I} - \mathbf{P}$. 
The derivation of equations (\ref{eqn:biasE}) and (\ref{eqn:varE}) is described in Appendix 1.
The expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$ and $\mathbf{C}_2$ are obtained using the results of Lemma \ref{lem:lemma1}.

@@

\begin{comment}
\section{STATISTICAL ANALYSIS - Ramp input}

### 2020-04-07 Extras/Summary.tex
Final version before internal defense

%!TEX root = ..\Thesis.tex
\chapter{Summary} \label{chap:Summary}

\vspace{-1.5cm}

Sensors are dynamical systems, and the physical quantities are their input excitations. 
The input magnitude must be estimated during the sensor transient state to expedite the measurements. 
Otherwise, users have to wait for the sensor steady state when the response is proportional to the input. 
In this case, the proportionality constant is the sensor static gain. 
One approach to get fast input estimation is filtering the sensor transient response with another dynamical system aiming to compensate the estimation time by regenerating the input. 
Another approach is processing the transient response on a digital signal processor (DSP). 
Contrary to model-based compensators, DSPs allow model-free methods that can reduce the measurement time.

There exists a data-driven method that directly estimates the step input level from the sensor step response. 
The formulation of this method is a Hankel-structure errors-in-variables (EIV) problem, where the regression matrix and the regressor are correlated.
The structured EIV problem is solved with recursive least-squares to allow for real-time implementation.
The range of application of this step input estimation method is broad because of its independence from the sensor model .

The stochastic properties of the step input estimation method are unknown and are not straightforwardly evident.
To validate the method for metrology applications, we studied the estimation uncertainty by doing a statistical analysis of the step input estimation.
The estimation bias and variance build the estimation uncertainty.
The statistical analysis of the estimation method yields expressions that predict the estimation first and second moments and enable the uncertainty assessment.

We computed the Cram\'er-Rao bound of the structured EIV problem to determine the minimum theoretical variance of its estimate.
Both in simulations and experiments, with temperature and weighing sensors, the sample mean squared error (MSE) is compared to the Cram\'er-Rao bound.
The simulation results show that the data-driven step input estimation is biased but with small variance, and that the estimation MSE is less than one order of magnitude larger than the Cram\'er-Rao bound.
In practical experiments, the step input estimation method is robust when the measurement noise is not Gaussian and white.

>>>>

%!TEX root = ..\Thesis.tex
\chapter{Summary} \label{chap:Summary}
\vspace{-1.5cm}
Measurements are dynamical processes where the input is estimated using the sensor response. 
The users cannot wait for the sensor to reach a steady regime when the sensor response closely approximates the sensor static gain times the input. 
To obtain a fast input estimation, one approach is filtering the sensor transient response with another dynamical system, aiming to compensate the estimation time.
The compensator is commonly built after a sensor model to perform deconvolution.  
Another approach is estimating the input with data-driven methods that are independent from the sensor model.
Current digital signal processors (DSP) allow the implementation of both model-based and data-driven estimation methods.

There exists a data-driven method that directly estimates the step input level from the sensor step response. 
The formulation of this method is a Hankel-structure errors-in-variables (EIV) problem, where the regression matrix and the regressor are correlated.
The structured EIV problem is solved with recursive least-squares for real-time implementation.
The independence from the sensor model of the estimation method broadens its range of application.

The uncertainty of the step input estimation method is unknown and is not straightforwardly evident.
To validate the method for metrology applications, we studied the stochastic properties of the step input estimation.
A statistical analysis yields the estimation bias and variance, and thus, the estimation uncertainty.

Both with simulations and experiments, the sample mean squared error (MSE) of the step input estimation method is compared to its analytical Cram\'er-Rao bound.
The Cram\'er-Rao bound determines the minimum theoretical variance of the estimation.
It was found that the data-driven step input estimation is biased but has small variance, and the estimation MSE is less than one order of magnitude larger than the Cram\'er-Rao bound.
Moreover, in practical experiments, the step input estimation method is robust when the measurement noise is not Gaussian and white.
